Reproducible Summary Tables with the gtsummary Package

### 2020-12-11 RJtemplate_gtsummary_submission.tex
Updates from Overleaf

\abstract{
An abstract of less than 150 words.
}

@@

@@

To show use of gtsummary functions, we will use a simulated clinical trial data set containing baseline characteristics of 200 patients who received Drug A or Drug B as well as the outcome of tumor response to the treatment.
Each variable in the data frame has been assigned an attribute label (i.e. \texttt{attr(trial\$trt, "label") == "Chemotherapy Treatment"}) with the labelled package\citep{labelled}. 
These labels are displayed in the {gtsummary} tables by default. Using {gtsummary} on a data frame without labels will print variable names in place of variable labels with the option to update the text displayed.

@@

% code for basic tbl_summary()
\begin{example}
trial %>%
  select(age, grade, response, trt) %>%
  tbl_summary(by = trt)
\end{example}
\begin{figure}[h!]
  \includegraphics[height=5cm]{summary_basic.png}
  \centering
\end{figure}

@@

\begin{example}
trial %>%
  select(age, grade, response, trt) %>%
  tbl_summary(
    by = trt,
    type = all_continuous() ~ "continuous2",
    label = age ~ "Patient Age",
    statistic = list(all_continuous() ~ c("{N_nonmiss}", 
                                          "{mean} ({sd})", 
                                          "{median} ({p25}, {p75})", 
                                          "{min}, {max}"),
                     all_categorical() ~ "{n} / {N} ({p}%)"),
    digits = all_categorical() ~ c(0, 0, 1),
    missing = "no"
  )
\end{example}
\begin{figure}[h!]
  \includegraphics[height=5cm]{summary_plus.png}
  \centering
\end{figure}

\textbf{A note about notation:}
Throughout the gtsummary package, you'll find function arguments that accept a list of formulas (or a single formula) as the input.
In the example above, the label for the age variable was updated using \texttt{label = age $\sim$ "Patient Age"}---equivalently, \texttt{label = list(age $\sim$ "Patient Age")}.
To select groups of variables, utilize the select helpers from the tidyselect\citep{tidyselect} and gtsummary.
Above, \texttt{all\_continuous()} was utilized to change the summary statistics for all continuous variables. 
Similarly, users may utilize \texttt{all\_categorical()} (from gtsummary), or any of the tidyselect helpers used throughout the tidyverse\citep{tidyverse} package, such as \texttt{starts\_with()}, \texttt{contains()}, etc.

@@

\end{example}
\begin{figure}[h!]
  \includegraphics[height=5cm]{summary_plus_plus.png}
  \centering
\end{figure}

@@

# summarize data
svy_api %>%
  tbl_svysummary(by = cname, 
                 include = c(growth, target, cname),
                 label = list(growth ~ "Growth",
                              target ~ "Target")) %>%
  add_p()
\end{example}
\begin{figure}[H]
  \includegraphics[height=4cm]{svysummary.png}
  \centering
\end{figure}

@@

\begin{example}
trial %>%
  tbl_cross(row = stage, col = trt, percent = "cell") %>%
  add_p(source_note = TRUE)
\end{example}
\begin{figure}[h!]
  \includegraphics[height=4cm]{cross.png}
  \centering
\end{figure}

@@

% TODO: I do not understand why this isn't being placed below the code! https://www.overleaf.com/learn/latex/Positioning_of_Figures
\begin{figure}[h!]
  \includegraphics[height=4cm]{survfit.png}
  \centering
\end{figure}

@@

The gtsummary package utilizes the gt package\citep{gt} to print the summary tables.
The gt package exports approximately one hundred functions to customize and style tables.
When you need to add additional details or styling not available within gtsummary, use the \texttt{as\_gt()} to convert the gtsummary object to gt and continue customization.

\begin{example}
trial %>%
  select(age, grade, trt) %>%
  tbl_summary(by = trt, missing = "no") %>%  
  add_overall() %>%    
  add_p() %>%      
  add_stat_label() %>%     
  modify_header(label ~ "**Variable**") %>%   
  modify_spanning_header(c(stat_1, stat_2) ~ "**Treatment Recieved**") %>%  
  as_gt() %>% 
  gt::tab_header(   
    title = gt::md("**Table 1. Patient Characteristics**"),
    subtitle = gt::md("_Highly Confidential_")
  ) %>%
  gt::tab_source_note("Data updated June 26, 2015")   
\end{example}
\begin{figure}[h!]
  \includegraphics[height=4cm]{custom.png}
  \centering
\end{figure}

@@

\begin{figure}[h!]
  \includegraphics[height=4cm]{regression.png}
  \centering
\end{figure}

The \texttt{tbl\_regression()} function leverages the huge effort behind the broom package's \texttt{tidy()} function to perform the initial formatting of the regression object\citep{broom}.
Because \texttt{tbl\_regression()} utilizes \texttt{broom::tidy()}, there are many model types that are supported out of the box, such as \texttt{lm()}, \texttt{glm()}, \texttt{lme4::lmer()}, \texttt{lme4::glmer()}, \texttt{geepack::geeglm()}, \texttt{survival::coxph()}, \texttt{survival::survreg()}, \texttt{survival::clogit()}, \texttt{nnet::multinom()}, \texttt{rstanarm::stan\_glm()}, models built with the mice package[add ref]and more. The gtsummary package also contains a handful of custom tidiers to report standardized coefficients and bootstrapped confidence intervals.

@@

\begin{figure}[h!]
  \includegraphics[height=4cm]{uvregression.png}
  \centering
\end{figure}

@@

The default styling (e.g. statistics displayed in \texttt{tbl\_summary()}, how p-values are rounded, decimal separator, and more) follow the reporting guidelines from European Urology, The Journal of Urology, Urology, and the British Journal of Urology International\citep{assel2019guidelines}.
However, you'll likely submit to another journal, or your personal preferences different from the defaults.
The gtsummary package is unique from other table building packages in the ability to set fine-grained customization defaults with themes. 
Themes were created to make these customizations easy to navigate and reuse across documents or projects. 
Several ready built themes are available in the package.
With themes, users can control default settings for existing functions (e.g. always present means instead of medians in \texttt{tbl\_summary()}), as well as other changes that are not modifiable with function arguments.

For example, using the theme for New England Journal of Medicine, large p-values are rounded to two decimal places and confidence intervals are shown as \texttt{"lb to ub"} instead of \texttt{"lb, ub"}.

@@

\begin{figure}[h!]
  \includegraphics[height=4cm]{nejm.png}
  \centering
\end{figure}

@@

\begin{figure}[h!]
  \includegraphics[height=4.5cm]{print_engines.png}
  \centering
\end{figure}

>>>>

\abstract{
The gtsummary package provides an elegant and flexible way to create publication-ready summary tables in R. The motivation behind the package stems from the work of statisticians, data scientists, and analysts, where every day, they summarize datasets and regression models in R, share results with collaborators, and eventually include them in published manuscripts or share with stakeholders. The gtsummary streamlines this process creating summaries of data sets, regression models, survey data, and survival data with simple code resulting in beautiful tables suitable for publication--often with a single line of code. The package follows a tidy framework, yielding an interface that is easy to learn that is also highly customizable.
}

@@

\citep{gtsummary}

@@

To show use of gtsummary functions, we will use a simulated clinical trial data set containing baseline characteristics of 200 patients who received Drug A or Drug B as well as the outcome of tumor response to the treatment.
Each variable in the data frame has been assigned an attribute label (i.e. \texttt{attr(trial\$trt, "label") == "Chemotherapy Treatment"}) with the labelled package \citep{labelled}. 
These labels are displayed in the {gtsummary} tables by default. Using {gtsummary} on a data frame without labels will print variable names in place of variable labels with the option to update the text displayed.

@@

% code for basic tbl_summary()
\begin{example}
trial %>%
  select(age, grade, response, trt) %>%
  tbl_summary(by = trt)
\end{example}
\begin{figure}[h!]
  \includegraphics[scale=0.28]{summary_basic.png}
  \centering
\end{figure}

@@

\begin{example}
trial %>%
  select(age, grade, response, trt) %>%
  tbl_summary(
    by = trt,
    type = all_continuous() ~ "continuous2",
    label = age ~ "Patient Age",
    statistic = list(all_continuous() ~ c("{N_nonmiss}", 
                                          "{mean} ({sd})", 
                                          "{median} ({p25}, {p75})", 
                                          "{min}, {max}"),
                     all_categorical() ~ "{n} / {N} ({p}%)"),
    digits = all_categorical() ~ c(0, 0, 1),
    missing = "no"
  )
\end{example}
\begin{figure}[h!]
  \includegraphics[scale=0.28]{summary_plus.png}
  \centering
\end{figure}

\textbf{A note about notation:}
Throughout the gtsummary package, you'll find function arguments that accept a list of formulas (or a single formula) as the input.
In the example above, the label for the age variable was updated using \texttt{label = age $\sim$ "Patient Age"}---equivalently, \texttt{label = list(age $\sim$ "Patient Age")}.
To select groups of variables, utilize the select helpers from the tidyselect \citep{tidyselect} and gtsummary.
Above, \texttt{all\_continuous()} was utilized to change the summary statistics for all continuous variables. 
Similarly, users may utilize \texttt{all\_categorical()} (from gtsummary), or any of the tidyselect helpers used throughout the tidyverse \citep{tidyverse} package, such as \texttt{starts\_with()}, \texttt{contains()}, etc.

@@

\end{example}
\begin{figure}[h!]
  \includegraphics[scale=0.28]{summary_plus_plus.png}
  \centering
\end{figure}

@@

# summarize data
svy_api %>%
  tbl_svysummary(by = cname, 
                 include = c(growth, target, cname),
                 label = list(growth ~ "Growth",
                              target ~ "Target")) %>%
  add_p()
\end{example}
\begin{figure}[h!]
  \includegraphics[scale=0.28]{svysummary.png}
  \centering
\end{figure}

@@

\begin{example}
trial %>%
  tbl_cross(row = stage, col = trt, percent = "cell") %>%
  add_p(source_note = TRUE)
\end{example}
\begin{figure}[h!]
  \includegraphics[scale=0.28]{cross.png}
  \centering
\end{figure}

@@

% TODO: I do not understand why this isn't being placed below the code! https://www.overleaf.com/learn/latex/Positioning_of_Figures
\begin{figure}[h!]
  \includegraphics[scale=0.28]{survfit.png}
  \centering
\end{figure}

@@

The gtsummary package utilizes the gt package \citep{gt} to print the summary tables.
The gt package exports approximately one hundred functions to customize and style tables.
When you need to add additional details or styling not available within gtsummary, use the \texttt{as\_gt()} to convert the gtsummary object to gt and continue customization.

\begin{example}
trial %>%
  select(age, grade, trt) %>%
  tbl_summary(by = trt, missing = "no") %>%  
  add_overall() %>%    
  add_p() %>%      
  add_stat_label() %>%     
  modify_header(label ~ "**Variable**") %>%   
  modify_spanning_header(c(stat_1, stat_2) ~ "**Treatment Recieved**") %>%  
  as_gt() %>% 
  gt::tab_header(   
    title = gt::md("**Table 1. Patient Characteristics**"),
    subtitle = gt::md("_Highly Confidential_")
  ) %>%
  gt::tab_source_note("Data updated June 26, 2015")   
\end{example}
\begin{figure}[h!]
  \includegraphics[scale=0.28]{custom.png}
  \centering
\end{figure}

@@

\begin{figure}[h!]
  \includegraphics[scale=0.28]{regression.png}
  \centering
\end{figure}

The \texttt{tbl\_regression()} function leverages the huge effort behind the broom package's \texttt{tidy()} function \citep{broom} to perform the initial formatting of the regression object.
Because \texttt{tbl\_regression()} utilizes \texttt{broom::tidy()}, there are many model types that are supported out of the box, such as \texttt{lm()}, \texttt{glm()}, \texttt{lme4::lmer()}, \texttt{lme4::glmer()}, \texttt{geepack::geeglm()}, \texttt{survival::coxph()}, \texttt{survival::survreg()}, \texttt{survival::clogit()}, \texttt{nnet::multinom()}, \texttt{rstanarm::stan\_glm()}, models built with the mice package[add ref]and more. The gtsummary package also contains a handful of custom tidiers to report standardized coefficients and bootstrapped confidence intervals.

@@

\begin{figure}[h!]
  \includegraphics[scale=0.28]{uvregression.png}
  \centering
\end{figure}

@@

The default styling (e.g. statistics displayed in \texttt{tbl\_summary()}, how p-values are rounded, decimal separator, and more) follow the reporting guidelines from European Urology, The Journal of Urology, Urology, and the British Journal of Urology International \citep{assel2019guidelines}.
However, you'll likely submit to another journal, or your personal preferences different from the defaults.
The gtsummary package is unique from other table building packages in the ability to set fine-grained customization defaults with themes. 
Themes were created to make these customizations easy to navigate and reuse across documents or projects. 
Several ready built themes are available in the package.
With themes, users can control default settings for existing functions (e.g. always present means instead of medians in \texttt{tbl\_summary()}), as well as other changes that are not modifiable with function arguments.

For example, using the theme for New England Journal of Medicine, large p-values are rounded to two decimal places and confidence intervals are shown as \texttt{"lb to ub"} instead of \texttt{"lb, ub"}. A theme was used to construct the tables shown in this manuscript to change the font to match the \emph{R Journal} font.

@@

\begin{figure}[h!]
  \includegraphics[scale=0.28]{nejm.png}
  \centering
\end{figure}

@@

\begin{figure}[h!]
  \includegraphics[scale=0.28]{print_engines.png}
  \centering
\end{figure}
